Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching

### 2025-11-24 sections/background_related.tex
missed a hat for x in iterative loop figures
Updates from Overleaf

\subsection{Formalization of Sound-Matching}
\label{sec:sound_matching_definition}
Following prior works~\cite{vahidi2023mesostructures,han2023perceptual}, we define sound-matching in terms of a parametric synthesizer, a target sound, a representation function, and a similarity measure. 
The key components are:
\begin{itemize}
    \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
    \item $x_\theta$: The synthesizer output, $x_\theta = g(\theta)$”.
    \item $\xzero$: The target sound to be replicated or imitated. 
    \item $\phi(\cdot)$: A representation (feature extraction) function mapping signals into a comparison space.
    \item $L$: A loss function that measures distance between $\xtheta$ and $\xzero$, typically via $L(\theta,\xzero) = d(\phi(\xtheta), \phi(\xzero))$ for some metric $d$.
\end{itemize}

@@

% Nodes
\node (start) [text centered] {\( g(\hat{\theta}) = x_{\theta} \)};
\node (L) [above of=start, right of=start, text centered] 
    {\( L(\hat{\theta}, x_{0}) \propto \phi(x_{\theta}) - \phi(x_{0}) \)};
\node (optimize) [below of=L, right of=L, text centered] {Optimize $\hat{\theta}$};
\node (new_theta) [below of=optimize, right of=start, text centered] {New \( \hat{\theta} \)};

@@

\subsection{Historical Framing of Sound-Matching}
Given the attributes discussed above, Table~\ref{tab:summary} summarizes relevant literature. Here we expand with a historical analysis of past sound-matching work.

@@

Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}. The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}. Possible chain functionalities are FM/AM, Low-Frequency Oscillators (\gls{LFO}), filters, and envelopes. Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.

 Audio embeddings can also be used as a similarity metric. In a recent work, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects. In their approach, a desired sound-effect is described in text and embedded using CLAP, followed by iterative updates from a gradient-free optimizer~\cite{evosax2022github} to the synthesizer's parameters to minimize embedding differences. Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}. However, it did yield better scores for ``artistic-interpretation'' (what we referred to as ``imitation''). It should be noted that audio embedding models are often trained using simpler loss functions---particularly those which utilize Fourier transforms---therefore improving such simpler loss functions is likely to improve the embedding models.

>>>>

\subsection{Formalization of Sound-Matching}
\label{sec:sound_matching_definition}
Following prior works~\cite{vahidi2023mesostructures,han2023perceptual}, we define sound-matching in terms of a parametric synthesizer, a target sound, a representation function, and a similarity measure. 
The key components are:
\begin{itemize}
    \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
    \item $x_\theta$: The synthesizer output, $x_\theta = g(\theta)$.
    \item $\xzero$: The target sound to be replicated or imitated. 
    \item $\phi(\cdot)$: A representation (feature extraction) function mapping signals into a comparison space.
    \item $L$: A loss function that measures distance between $\xtheta$ and $\xzero$, typically via $L(\theta,\xzero) = d(\phi(\xtheta), \phi(\xzero))$ for some metric $d$.
\end{itemize}

@@

% Nodes
\node (start) [text centered] {\( g(\hat{\theta}) = x_{\hat{\theta}} \)};
\node (L) [above of=start, right of=start, text centered] 
    {\( L(\hat{\theta}, x_{0}) \propto \phi(x_{\hat{\theta}}) - \phi(x_{0}) \)};
\node (optimize) [below of=L, right of=L, text centered] {Optimize $\hat{\theta}$};
\node (new_theta) [below of=optimize, right of=start, text centered] {New \( \hat{\theta} \)};

@@

\subsection{Historical Framing of Sound-Matching}
Table~\ref{tab:summary} summarizes relevant literature. Following the discussion of important sound-matching components, a historical analysis of past work can be given.

@@

Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects along with a loss function that combines P-Loss with a \textit{signal-chain loss}. The synthesizer is a customizable chain of effects, which feeds one output as input to the next step of the chain; signal-chain loss compares the differences in parameter and output at every output step in the chain~\cite{uzrad2024diffmoog}. Possible chain functionalities are FM/AM, low-frequency oscillators (\gls{LFOs}), filters, and envelopes. Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.

Audio embeddings can also be used as a similarity metric. For example, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects. In their approach, a desired sound-effect is described in text and embedded using CLAP, followed by iterative updates from a gradient-free optimizer~\cite{evosax2022github} to the synthesizer's parameters to minimize embedding differences. Based on manual hearing tests, this approach did not produce the ``correct'' sounds more often than prior methods~\cite{kreuk2022audiogen}. However, it did yield better scores for ``artistic-interpretation'' (what we referred to as ``imitation''). It should be noted that audio embedding models are often trained using simpler (often Fourier-based) loss functions, therefore improvements to the fundamental loss functions may also lead to better embeddings.
